\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2017-05-02 redaction/dhp_sept2016.tex
Pour soumission DEGIT

\documentclass[a4paper,11pt]{article}
%\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
%\usepackage[latin1]{inputenc}
%\usepackage{hyperref}
%\usepackage[T1]{fontenc}
%\usepackage[utf8]{inputenc}
%%GD Quand je mets usepackage[utf8]{inputenc} je n'arrive plus à compiler la biblio
%J'ai essayé de débuger : je n'y suis pas arrivé
\usepackage{rotating}
\usepackage{setspace}
\usepackage{lscape}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{vmargin}
\usepackage{epstopdf}
\usepackage{hyperref}
\usepackage{float}
\usepackage{caption}
\usepackage[tableposition=top]{caption}
\usepackage{ bbold }
\usepackage{ upgreek }
\usepackage{comment}
\includecomment{commentGD}
\usepackage[draft]{todonotes}
%%GDPour cacher les notes, \usepackage[disable]{todonotes}

@@

%\todo{reprendre slides presentation LEDa pour la motivation de la question.}
%JH: apparently, it has been done! Or do you want to have the Italian shoes example in the intro????

@@

Coming now to interpretation, Table \ref{tab:summary_results} shows the magnitude of overall transport costs is sizeable. Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping. Furthermore, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable. Over the whole period, additive costs are 2.9\% of the export price for ocean shipping  - and ad-valorem ones being equal to 3.2\%. These values are respectively equal to 1.8\% and 2.5\% for air transport. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward. In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.%\todo{Ajouter un commentaire sur mean different de median, en comparaison avec les data.}

@@

%\todo{Comparer avec la litterature. Cf AvW 2004 pure transport costs 11\% markup over production costs.}

@@

\textbf{a dire pas forcement ici, tout n'est pas que sur cette question des effets de composition.}

\textbf{argument general}
In comparison with \citet{hummels2007}, our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs. Precisely, we allow it to change over time and across country partners and products, while \citet{hummels2007} imposes it to be constant in these three dimensions.\footnote{In the terms of \cite{hummels_skiba}, our estimation strategy allows for a varying $\beta$ across time, country partner and product.}

\textbf{argument sur les effets de composition}
If we share this view with \citet{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects. First, we provide a broader characterization of the time trends in transport costs as our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs. In characterizing the time trend in the ``pure transport costs'', \citet{hummels2007} eliminates the changes in the price per kilo dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p. 146.} By contrast, we provide an explicit decomposition of this additive cost component over time: By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we are thus able to characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time. As well, we are able to do so for the multiplicative component of trade costs. We also agglomerate the two components in a unified measure of transport costs from which we exclude the composition effects. This allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones, reported in Figure \ref{fig:decomp_TC_3d}). Last, for sake of comparison though, we do the same exercise as \citet{hummels2007}.

\textbf{A faire: sous forme de graph, par transport mode : Ligne 1, l'exercice de Hummels, ln TC in \% fob price (observed)= dummy pays-produit + beta* ln (weight/value) + dummy temps. Ligne 2, ce qu'on a sur la dimension iceberg cost estimated; ligne 3, sur les couts additifs estimes. Ligne 4, a faire, la somme des deux, estimee? Ou alors, partir simplement du TC observe, ie ce qu'on a deja?}

Second, we characterize the evolution of the ``pure'' transport costs over time by measuring how these costs have changed over time, blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample), while \citet{hummels2007} captures the trend evolution of by the yearly dummy. Doing so enables us to build measures of transport costs that are easily comparable between transport modes and transport costs components. \textbf{Pertinence de cet argument? A reformuler.}

@@

\textbf{Faire aussi une analyse plus fine par grande categorie: Manuf / Agriculture / Matieres premieres.}

@@

\paragraph{what else?}

>>>>

\documentclass[a4paper,11pt]{article}
%\usepackage[pdftex]{graphicx}
\usepackage{amsmath}
%\usepackage[latin1]{inputenc}
%\usepackage{hyperref}
%\usepackage[T1]{fontenc}
%\usepackage[utf8]{inputenc}
%%GD Quand je mets usepackage[utf8]{inputenc} je n'arrive plus à compiler la biblio
%J'ai essayé de débuger : je n'y suis pas arrivé
\usepackage{rotating}
\usepackage{setspace}
\usepackage{lscape}
\usepackage[round]{natbib}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{vmargin}
\usepackage{epstopdf}
\usepackage{hyperref}
\usepackage{float}
\usepackage{caption}
\usepackage[tableposition=top]{caption}
\usepackage{ bbold }
\usepackage{ upgreek }
\usepackage{comment}
\includecomment{commentGD}
\usepackage[draft]{todonotes}
%\usepackage[disable]{todonotes}
%%GDPour cacher les notes, \usepackage[disable]{todonotes}

@@

\todo{reprendre slides presentation LEDa pour la motivation de la question.}
\todo{JH: apparently, it has been done! Or do you want to have the Italian shoes example in the intro????}

@@

Coming now to interpretation, Table \ref{tab:summary_results} shows the magnitude of overall transport costs is sizeable. Over 1974-2013, they correspond to a mean increase of the export price by 5.8\% for ocean shipping, and by 5.1\% for air shipping. Furthermore, when decomposing transport costs into an additive and an ad-valorem component, we find that the per-kg cost dimension is sizeable. Over the whole period, additive costs are 2.9\% of the export price for ocean shipping  - and ad-valorem ones being equal to 3.2\%. These values are respectively equal to 1.8\% and 2.5\% for air transport. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward. In quantitative terms, the omission of the additive term leads to overestimate the ad-valorem component by roughly a factor 2.\todo{Ajouter un commentaire sur mean different de median, en comparaison avec les data.}

@@

\todo{Comparer avec la litterature. Cf AvW 2004 pure transport costs 11\% markup over production costs.}

@@

\todo{a dire pas forcement ici, tout n'est pas que sur cette question des effets de composition.}

\todo{argument general}
In comparison with \citet{hummels2007}, our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs. Precisely, we allow it to change over time and across country partners and products, while \citet{hummels2007} imposes it to be constant in these three dimensions.\footnote{In the terms of \cite{hummels_skiba}, our estimation strategy allows for a varying $\beta$ across time, country partner and product.}

\todo{argument sur les effets de composition}
If we share this view with \citet{hummels2007}, our strategy for eliminating composition effects differs from his in two main respects. First, we provide a broader characterization of the time trends in transport costs as our empirical analysis allows for a more flexible estimation of the share between the additive and the ad-valorem components of transport costs. In characterizing the time trend in the ``pure transport costs'', \citet{hummels2007} eliminates the changes in the price per kilo dimension, attributing this to a change in the composition of the bundle over time (per country-commodity).\footnote{See \citet{hummels2007}, p. 146.} By contrast, we provide an explicit decomposition of this additive cost component over time: By comparing the additive cost component unfitted (total) and fitted (composition effects excluded), we are thus able to characterize if this dimension of the transport cost has fallen due to (for instance) the fact that product quality has increased over time (weight has reduced, hence reducing the transport cost by dollar transported), or if it is the ``pure'' additive costs (for instance, handling costs) that have reduced over time. As well, we are able to do so for the multiplicative component of trade costs. We also agglomerate the two components in a unified measure of transport costs from which we exclude the composition effects. This allows us to compare how ``pure'' transport costs have changed over time, composition effects excluded, and to compare them to the unfitted ones, reported in Figure \ref{fig:decomp_TC_3d}). Last, for sake of comparison though, we do the same exercise as \citet{hummels2007}.

\todo{A faire: sous forme de graph, par transport mode : Ligne 1, l'exercice de Hummels, ln TC in \% fob price (observed)= dummy pays-produit + beta* ln (weight/value) + dummy temps. Ligne 2, ce qu'on a sur la dimension iceberg cost estimated; ligne 3, sur les couts additifs estimes. Ligne 4, a faire, la somme des deux, estimee? Ou alors, partir simplement du TC observe, ie ce qu'on a deja?}

Second, we characterize the evolution of the ``pure'' transport costs over time by measuring how these costs have changed over time, blocking the composition of trade flows by product and country partners to the one observed in 1974 (the beginning of our sample), while \citet{hummels2007} captures the trend evolution of by the yearly dummy. Doing so enables us to build measures of transport costs that are easily comparable between transport modes and transport costs components. \todo{Pertinence de cet argument? A reformuler.}

@@

\todo{Faire aussi une analyse plus fine par grande categorie: Manuf / Agriculture / Matieres premieres.}

@@

\todo{what else?}
